Clouds and Radiation in a mock-Walker Circulation

### 2019-10-16 Silvers_WalkerCell.tex
several changes to Silvers_WalkerCell.tex, mostly textual changes

  useful tool with which to explore the basic physics of climate and climate sensitivity, but they lack a large-scale circulation.}

@@

  in particular.}

@@

The mock-Walker Circulation that emerges from these simulations is shown in Figures \ref{fig:rh_psi_P25vsE25} and 
\ref{fig:precip_vertvel} to be characterized by a strong overturning circulation with precipitation focused over the warmer 
SSTs and a humid boundary layer across the full length of the domain.   For the GCM-like configuration (P25, left panels 
of Figure \ref{fig:rh_psi_P25vsE25}) with fully parameterized convection and with the longwave CRE turned on, there is 
a strong overturning circulation that results in a humid, condensate loaded troposphere near the middle of the domain 
and a dry ($< 10\% RH$) subsidence region with almost no condensate above $800 \rm{hPa}$.    
To illustrate how (how what?  the double-plume convective parameterization influences the atmospheric state) 
we compare the P25 experiment with analogous experiments in which the longwave CRE is off (P25 LWCRE off, middle 
panels of Figure \ref{fig:rh_psi_P25vsE25}) and in which the convection is made explicit by turning off the convective 
parameterization (E25, right panels of Figure \ref{fig:rh_psi_P25vsE25}).  
The mass streamfunction and vertical velocity both show the P25 experiment has a stronger overturning circulation 
than either E25 or P25 LWCRE off.  P25 also has higher humidity in the deep convective region, and lower humidity in 
the subsidence region over the 
cooler SST (Figure \ref{fig:rh_psi_P25vsE25}) than either E25 or P25 LWCRE off.  Averaged over the full domain, 
the parameterized convection leads to a dryer atmosphere with less 
condensate (both liquid and ice) and with weaker radiative cooling and condensational heating (below 800 hPa).  However, in 
the updraft region the convective parameterization results in more moisture aloft.

@@

One of the most prominent features of our GCM-like (in particular, P25) simulations is an asymmetry in both the time-dependent and 
steady-state solutions (Figures \ref{fig:rh_psi_P25vsE25},  \ref{fig:precip_vertvel}, \ref{fig:domdep} and  \ref{fig:conv_vs_ls}).  The steady-state precipitation maximum is located  not over the warmest SST but is
shifted to slightly cooler temperatures.  The asymmetry is apparent in the vertical velocity, mass circulation, relative humidity, 
specific humidity, and radiative heating.  
For most of these cases the asymmetry appears as an aversion of the precipitation maxima to reside over the SST maximum.  
For the P25 case shown in Figure \ref{fig:rh_psi_P25vsE25} a strong ($1 \, \rm{m \, s^-1}$) domain
mean shear develops above about 500 \textit{m} which shifts the precipitation and circulation off center for years at a time.
When the convective parameterization
is turned off (E25), the overturning circulation becomes weaker and broader (w, and psi), and the precipitation, cloud fields, and 
circulation all consistently reside over the SST maximum.   
While the parameterized convection plays a large role in driving this asymmetry it appears to be not entirely a 
result of the convective parameterization, but also due to an interaction between the convective 
parameterization and the LW CRE.   The degree to which this asymmetry influences the comparison with other experiments is 
unclear.    Note: check the wind shear for P25large, P100small, and P100large (E25 has less shear than E1 or E2).

@@

The abnormality, or asymmetry, of the P25 simulation was the first surprising feature noticed by the authors.
This asymmetry is also present on larger domains and at a resolution of $100 \, \rm{km}$.  For some of the other cases observed, such
as P100 (on the small domain), the asymmetry can be described as an aversion of the precipitation maxima to reside over the 
SST maximum while in other cases, such as the P25 case shown in \ref{fig:rh_psi_P25vsE25}, a strong ($1 \, \rm{m s^-1}$) domain
mean shear develops which shifts the precipitation and circulation off center for years at a time.  This appears to be not entirely a 
result of the convective parameterization, but due to an interaction between the convective parameterization and the LW CRE.

@@

Now talk about the organizing influence of LWCRE...

@@

Because of the close connection between the domain mean value of precipitation, atmospheric condensational heating, and 
the total radiative cooling the time evolution of the precipitation serves as a useful measure of whether a model has 
reached statistical equilibrium.  Figure \ref{fig:precip_dom_mn} demonstrates that after about 30 days, the P25, E2, and E1
simulations begin to oscillate about a mean state that is roughly constant over periods greater than about 30 days. 
The large oscillations in domain mean precipitation are similar to those noted in previous studies 
(Silvers et al., 2016; Patrizio and Randall, 2019).   The experiments with a higher resolution have a lower value of domain
mean precipitation, with the E2 simulation having the lowest (3.1 mm/day).  This can be understood as a consequence
of the differences in upper level cloud fraction and will be discussed further in a later section.

%One of the most prominent features of our 25km grid-spacing simulations is an asymmetry in both the time-dependent and 
%steady-state solutions.  This results in the steady-state precipitation maximum locating off the maximum in SST and is apparent in 
%the vertical velocity, mass circulation, relative humidity, specific humidity, and radiative heating.   A tendency of the precipitation to 
%occur off of the SST maximum is also apparent in the simulations with a grid-spacing of 100km, although the 25km simulations display 
%an irregularity that is greater than the 100km simulations.  Complex patterns of precipitation over a fixed sinusoidal or Gaussian SST 
%distribution has been noted many times 
%(Bretherton et al. 2006, Kuang and Wofsy, Ming Zhao (personal communication, see Jeevanjee et al. 2018)) but the irregularities have 
%tended to be symmetric, unlike our results.   
%
%For the simulations that are closest to the set up of a GCM, with fully parameterized convection, the steady-state circulation and 
%precipitation maximum are located not over the warmest SST but shifted to regions about one degree cooler.  For the case
%with a grid-spacing of 25km and a domain width of 4000km, the precipitation maximum can remain shifted off of the maximum in
%SST for multiple years.   While focused within the central quarter of the domain, simulations on larger domains and with a 
%grid-spacing of 100km also show an aversion to settling over the SST maximum.   When the convective parameterization
%is turned off, the overturning circulation becomes weaker and broader (w, and psi), but the precipitation, cloud fields, and 
%circulation all consistently reside over the SST maximum.  When the convective parameterization is turned on, a domain mean 
%wind shear develops throughout the lowest 10km that acts to shift the convective activity and circulation off of the SST maximum.  
%Note: check the wind shear for P25large, P100small, and P100large.

%% low level structure
%\begin{figure}
%\begin{center}
%\includegraphics[width=0.95\columnwidth]{walkerfigs/lowlevel_therm_test.eps}
%\caption{{Low-level thermodynamic structure (a-e) and vertical shear (lower right, f).  Solid lines show experiments that include LWCRE, 
%dashed lines experiments without LWCRE.  The yellow dashed-dot line is a 25 km experiment without LWCRE and with all 
%convective parameterization turned off.   Panel e shows the temperature at the lowest model level (thin lines) and the virtual 
%temperature at the lowest model level(thick lines).  The thick black line is the prescribed surface temperature.}}
%\end{center}
%\end{figure}
%

@@

%%%%
% section 2 of results

%% PDF for location of precip maximum of precipitation for experiments at both 25km and 100km resolutions 
%% LW CRE off
%\begin{figure}
%  \centering
%      %\includegraphics[width=0.96\columnwidth]{walkerfigs/hist_4pan_25vs100_lwcreoff.eps}
%      \includegraphics[width=0.6\columnwidth]{walkerfigs/hist_4pan_25vs100_lwcre_draft.eps}
%  \caption{Normalized probability distributions of the location of the maximum value of precipitation for P25 and P100
%  experiments .   Panels on the right (left) show experiments on the domain with a long dimension of 4000 km 
%  (16000 km).  Black distributions are for experiments with the LW CRE on and green distributions for experiments 
%  with LW CRE off.  The black curves show a Gaussian function with a standard deviation computed for 
%  each experiment and an expected location in the center of the domain where the SST is warmest.}
%\end{figure}
%

When designing numerical experiments the size of the domain and the grid-spacing that determines model resolution are two of the most 
fundamental choices that must be made.   One the goals of this study was to explore the results of changing grid-spacing over a wide 
range of values.  To simplify the analysis we have in most cases chosen to keep the 
long horizontal dimension fixed at $4000\, \rm{km}$.  However, previous studies (e.g. Bretherton et al., 2006; Muller and Held, 2012; Jeevanjee and Romps, 2013) have 
documented sensitivities to domain size.   The analysis of the previous, and of the next section focus on results from experiments using
a domain with a long domain length of $4000 \, \rm{km}$.  However we do find sensitivity to the domain size and thought it to be
sufficiently interesting to report.

Examining the evolution in time of the precipitation field reveals important characteristics of these simulations, and 
illustrates how much the distribution can vary as a function of domain size, parameterization of convection, and the effect of the 
longwave radiation due to clouds. 
Shown in Figures {\ref{fig:domdep}} and {\ref{fig:domdep_lwoff}} are Hovmoller plots of precipitation after 
averaging along the short horizontal dimension.  The four panels show simulations with a grid-
spacing of $25\, \rm{km}$ on a grid with a horizontal width of $4000\, \rm{km}$ (far right) and $16000\, \rm{km}$ (middle right), and 
simulations with a grid-spacing of $100\, \rm{km}$ on a grid with a horizontal width of $4000\, \rm{km}$ (middle left) and 
$16000\, \rm{km}$ (far left).  Figure  {\ref{fig:domdep_lwoff}} shows the equivalent simulations with LW CRE off.    
 At all resolutions the Hovmoller plots show that the longwave CRE acts to concentrate the 
precipitation over a smaller geographic extent.   This is consistent with the idea that CREs strengthen an overturning circulation and
narrow the region of deep convection (see Popp and Silvers, 2017, HH, Dixit et al. 2019.  etc.).
The structure of the precipitation changes more as a function of domain size than it does as a function of grid-spacing. 
On the large domains, the difference between experiments with and without LW CRE is extreme.  In contrast to the control 
experiments in Figure {\ref{fig:domdep}} which all show a narrow region of strong precipitation which meanders within 
500 km of the SST maximum at the center of the domain,  
the experiments without the LW CRE basically have no region with consistently strong precipitaiton.  Instead, 
smaller cells or lines of precipitation develop within an area that is roughly 8000 km wide.

@@

Smaller domain lengths were found to have a more focused ascent region and larger precipitation rates in Bretherton et al., 2006.
They also found less low-level clouds over the colder SSTs in the small domains, relative to a control domain size that was 4x
larger.   We have found larger precipitation rates for the larger domains.  Does our larger domain ascent region change?

@@

Although the domain mean total precipitation is constrained by the radiative cooling of the atmosphere, the amount of 
precipitation that is produced by the convection scheme or the large-scale cloud scheme is not constrained. 
The distribution of convective and large-scale precipitation can often give a hint of how the condensational heating
in a GCM is being distributed among the parameterizations.  Precipitation from each of these two components is
shown in Figure \ref{fig:conv_vs_ls}.  From this we can see that in the regions of large-scale ascent, most of the 
precipitation derives from the large-scale cloud scheme.  We also see that the LW CRE strongly influences the 
large-scale precipitation but has a much weaker impact on the convective precipitation.  With the single exception 
of the P100 LW CRE off experiment, the convective precipitation
produces relatively little of the total precipitation for both resolutions and on both domains.

@@

\section{The mock-Walker Circulation Dependence on Grid-Spacing: A transition from a 
General Circulation Model to a Cloud Resolving Model}

>>>>

  useful tool with which to explore the basic physics of climate and climate sensitivity, but they lack a large-scale circulation.  Many 
  studies have also shown the importance and prevalence of convective self aggregation (CSA).  A mock-walker circulation with an
  imposed SST gradient can be thought of as studying a partially constrained aggregating or organized system in a way that removes
  some of the ambiguity.  The mock-Walker circulation also allows for a simple way of studying in physical space, the connection
  between column relative humidity and circulation in a way that is directly tied to the physical state of the system rather than with
  a forced compositing of the data (Bretherton et al. 2005, Muller and Held, 2012).}

@@

  in particular.  To what extent do RCE simulations of CSA agree between GCMs and CRMs?  Perhaps here I can hint at the utility of 
  this methodology for unifying our understanding.}

@@

The mock-Walker Circulation that emerges from these simulations is shown in Figures \ref{fig:rh_psi_P25vsE25} and 
\ref{fig:precip_vertvel} to be characterized by a strong overturning circulation with precipitation focused over the warmer 
SSTs and a humid boundary layer across the full length of the domain.   For the GCM-like configuration (P25, left panels 
of Figure \ref{fig:rh_psi_P25vsE25}) with fully parameterized convection and with the longwave CRE turned on, there is 
a strong overturning circulation that results in a humid, condensate loaded troposphere near the middle of the domain 
and a dry ($< 10\% RH$) subsidence region with almost no condensate above $800 \rm{hPa}$.    
To illustrate how (how what?  the double-plume convective parameterization influences the atmospheric state) 
we compare the P25 experiment with analogous experiments in which the longwave CRE is turned off (P25 LWCRE off, middle 
panels of Figure \ref{fig:rh_psi_P25vsE25}) and in which the convection is made explicit by turning off the convective 
parameterization (E25, right panels of Figure \ref{fig:rh_psi_P25vsE25}).  
The mass streamfunction and vertical velocity both show the P25 experiment has a stronger overturning circulation 
than either E25 or P25 LWCRE off.  P25 also has higher humidity in the deep convective region, and lower humidity in 
the subsidence region over the 
cooler SST (Figure \ref{fig:rh_psi_P25vsE25}) than either E25 or P25 LWCRE off.  Averaged over the full domain, 
the parameterized convection leads to a dryer atmosphere with less 
condensate (both liquid and ice) and with weaker radiative cooling and condensational heating (below 800 hPa).  However, in 
the updraft region the convective parameterization results in more moisture aloft.

@@

One of the most prominent features of our GCM-like (in particular, P25) simulations is an asymmetry relative to the 
symmetric SST distribution in both the time-dependent and 
steady-state solutions (Figures \ref{fig:rh_psi_P25vsE25},  \ref{fig:precip_vertvel}, \ref{fig:domdep} and  \ref{fig:conv_vs_ls}).  
This asymmetry is also present on larger domains and at a resolution of $100 \, \rm{km}$. 
The steady-state precipitation maximum is located  not over the warmest SST but is
shifted to slightly cooler temperatures.  The asymmetry is apparent in the vertical velocity, mass circulation, relative humidity, 
specific humidity, and radiative heating.  
For most of these cases the asymmetry appears as an aversion of the precipitation maxima to reside over the SST maximum.  
For the P25 case shown in Figure \ref{fig:rh_psi_P25vsE25} a strong ($1 \, \rm{m \, s^-1}$) domain
mean shear develops above about 500 \textit{m} which shifts the precipitation and circulation off center for years at a time.
When the convective parameterization is turned off (E25), the overturning circulation becomes weaker and 
broader (w, and psi), and the precipitation, cloud fields, and 
circulation all consistently reside over the SST maximum.   
While the parameterized convection plays a large role in driving this asymmetry it appears to be not entirely a 
result of the convective parameterization, but also due to an interaction between the convective 
parameterization and the LW CRE.   The degree to which this asymmetry influences the comparison with other experiments is 
unclear.    Note: check the wind shear for P25large, P100small, and P100large (E25 has less shear than E1 or E2).

@@

%The abnormality, or asymmetry, of the P25 simulation was the first surprising feature noticed by the authors.
%This asymmetry is also present on larger domains and at a resolution of $100 \, \rm{km}$.  For some of the other cases observed, such
%as P100 (on the small domain), the asymmetry can be described as an aversion of the precipitation maxima to reside over the 
%SST maximum while in other cases, such as the P25 case shown in \ref{fig:rh_psi_P25vsE25}, a strong ($1 \, \rm{m s^-1}$) domain
%mean shear develops which shifts the precipitation and circulation off center for years at a time.

@@

The default experimental configuration includes radiation that is fully interactive with the clouds, water vapor, and 
temperature of the atmospheric state.  It can be seen in the left panels of Figure \ref{fig:rh_psi_P25vsE25} that this 
results in a stronger circulation with more condensate in the deep convective region and a dryer mid-troposphere
relative to the case with the LW CRE off.   A stronger and spatially concentrated circulation for cases when the 
clouds interact with the longwave radiation can also be seen in Figures 4, 5, 9 and 10 and is particular apparent in 
the cases with a large domain (16000 km long dimension).  This will be discussed further in the next section.

@@

Because of the tight constraints that connect the domain mean precipitation, atmospheric condensational heating, and 
the total radiative cooling, the time evolution of the precipitation serves as a useful measure of whether a model has 
reached statistical equilibrium.  Figure \ref{fig:precip_dom_mn} demonstrates that this equilibrium is reached after about 30 
days for the E2, and E1 simulations, and after about 50 days for the P25 and P100 simulations.   After the initial adjustments
the simulations all oscillate about mean precipitation values which, with the exception of E2, increases with the grid-spacing 
(Table 1).   The fact that E1 and E2 reach equilibrium sooner than E25, P25, or P100 and that the period of oscillation 
about the domain mean precipitation is smaller helps to justify the 6 month simulation times for E1 and E2.  
%The period of oscillation is much larger for the GCM like simulations than it is for the CRM like oscillations.
The large oscillations in domain mean precipitation are similar to those noted in previous studies 
(Silvers et al., 2016; Patrizio and Randall, 2019).   
%The experiments with a higher resolution have a lower value of domain
%mean precipitation, with the E2 simulation having the lowest (3.1 mm/day).  
This can be understood as a consequence
of the differences in upper level cloud fraction and the surface energy budget 
and will be discussed further in a later section.

@@

When designing numerical experiments the size of the domain and the grid-spacing that determines model resolution are 
two of the most fundamental choices that must be made.   One the goals of this study was to explore the results of 
changing grid-spacing over a wide range of values.  To simplify the analysis we have in most cases chosen to keep the 
long horizontal dimension fixed at $4000\, \rm{km}$.  However, previous studies 
(e.g. Bretherton et al., 2006; Muller and Held, 2012; Jeevanjee and Romps, 2013) have documented sensitivities to domain 
size.   The analysis of the previous, and of the next section focus on results from experiments using
a domain with a long domain length of $4000 \, \rm{km}$.  However we do find interesting sensitivities to the domain size that are noted here.

The evolution in time of the precipitation field reveals important characteristics of these simulations, and 
illustrates how much the distribution can vary as a function of domain size, parameterization of convection, 
and the effect of the longwave radiation due to clouds. 
Shown in Figures {\ref{fig:domdep}} and {\ref{fig:domdep_lwoff}} are Hovmoller plots of precipitation after 
averaging along the short horizontal dimension.  The four panels show simulations with a grid-
spacing of $25\, \rm{km}$ on a grid with a horizontal width of $4000\, \rm{km}$ (far right) and $16000\, \rm{km}$ (middle right), and 
simulations with a grid-spacing of $100\, \rm{km}$ on a grid with a horizontal width of $4000\, \rm{km}$ (middle left) and 
$16000\, \rm{km}$ (far left).  Figure  {\ref{fig:domdep_lwoff}} shows the equivalent simulations with LW CRE off.    
 At all resolutions the Hovmoller plots show that the longwave CRE acts to concentrate the 
precipitation over a smaller geographic extent.   This is consistent with the idea that CREs strengthen an overturning circulation and
narrow the region of deep convection (see Popp and Silvers, 2017, HH, Dixit et al. 2019.  etc.).
The structure of the precipitation changes more as a function of domain size than it does as a function of grid-spacing. 
On the large domains, the difference between experiments with and without LW CRE is extreme.  In contrast to the control 
experiments in Figure {\ref{fig:domdep}} which all show a narrow region of strong precipitation which meanders within 
500 km of the SST maximum at the center of the domain,  
the experiments without the LW CRE basically have no region with consistently strong precipitation.  Instead, 
smaller cells or lines of precipitation develop within an area that is roughly 8000 km wide.  As previously notes, these
GCM-like simulations equilibrate after approximately 50 days.  However, it is interesting to note a dramatic change in 
the distribution of precipitation on the $4000\, \rm{km}$ domain simulations after almost 2 years.   The domain mean 
precipitation does not significantly change in these cases, only the spatial structure.

@@

Smaller domain lengths were found to have a more focused ascent region and larger precipitation rates in Bretherton et al., 2006.
They also found less low-level clouds over the colder SSTs in the small domains, relative to a control domain size that was 4x
larger.   We have found larger precipitation rates for the larger domains.  Based on the precipitation fields (Hovmollers) it is 
possible that we have fewer low-level clouds over the colder SSTs in agreement with B06, but this needs to be checked.   Does our larger domain ascent region change?

@@

Although the domain mean total precipitation is constrained by the radiative cooling of the atmosphere, the amount of 
precipitation that is produced by the convection scheme or the large-scale cloud scheme is not constrained. 
The distribution of convective and large-scale precipitation can often indicate how the condensational heating
in a GCM is being distributed among the parameterizations.  Precipitation from each of these two components is
shown in Figure \ref{fig:conv_vs_ls}.  From this we can see that in the regions of large-scale ascent, most of the 
precipitation derives from the large-scale cloud scheme.  We also see that the LW CRE dramatically increases
the large-scale precipitation.  The LW CRE has a much smaller effect on the magnitude of the convective precipitation 
but acts to spatially concentrate it.  With the single exception 
of the P100 LW CRE off experiment, the convective precipitation produces relatively little of the total precipitation for 
both resolutions and on both domains.

@@

%\section{The mock-Walker Circulation Dependence on Grid-Spacing: A transition from a 
%General Circulation Model to a Cloud Resolving Model}
\section{Dependence on Resolution: From a General Circulation Model to a Cloud Resolving Model}
